QECO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2024-10-09 Revision/Paper_Rev.tex
Rev_V5.6

	\item \textit{Performance Evaluation:} \textcolor{blue}{We conduct comprehensive experiments to evaluate the QECO's performance under various computational workloads, as well as its training convergence with different techniques and hyperparameters. The results demonstrate that our algorithm effectively utilizes the computing resources of MDs and ENs, addressing the dynamic workloads at ENs. It converges more quickly compared to DQN and D3QN methods, while providing a substantial improvement of at least 34.8\% in average QoE. This advantage is achieved through prioritizing MD's QoE preferences, resulting in improvements of at least 12.2\%, 11.7\%, and 10.8\% in the average number of completed tasks, task delay, and energy consumption, respectively, when compared to the distributed and collective DRL-based offloading algorithm (DCDRL) \cite{qiu2020distributed} and several benchmark methods.}

@@

	  \textcolor{blue}{Some studies explored task offloading problems in more advanced scenarios. Chen \textit{et al.} in \cite{chen2021drl} addressed a joint optimization problem involving computation offloading and resource allocation, aiming to reduce both task processing delay and energy consumption across all MDs.
	 To optimize privacy protection and quality of service, authors in \cite{wu2024privacy} investigated the joint computation offloading and power allocation problems for the IIoT network. They modeled the problem as an MDP and proposed a multi-agent deep Q-network (DQN)-based algorithm.
	 Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm to concurrently optimize resource allocation and computation offloading.
	 In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
	 These works did not take into account the challenges of a dynamic network environment. For instance, in \cite{chen2021drl}, MDs are assumed to have sufficient transmission and computation resources from ENs, which may not always align with real-world conditions. While the works of \cite{wu2024privacy}--\cite{zhao2019deep} investigated the limited resources in MEC, they did not consider the dynamic workload at ENs in decision making processes, imposing significant challenges in overall system performance. }

@@

	 \begin{table*}[htbp]\textcolor{blue}{
		\renewcommand{\arraystretch}{0.7}
		\captionsetup{name=TABLE}
		\caption{Comparison of Related Works Characteristic}
		\scalebox{1}{%
			\begin{tabular}{ lp{2.2cm}p{3cm}p{3.4cm}p{0.7cm}p{2cm}p{3.2cm}l} 
				\toprule
				\textbf{Paper} & \textbf{Scenario} &  \textbf{Problem} &  \textbf{Objective} &  \textbf{Model} & \textbf{Method} &  \textbf{Drawbacks}  \\ \midrule
				\cite{liao2023online} & Queuing-based MEC & Online computation offloading & Minimize computation delay and energy consumption & MDP & DDPG + DDQN & Only for a singel MEC system model  \\\midrule
				\cite{zhou2021deep} & heterogeneous MEC & Resource allocation and computation offloading & Minimize energy consumption of the entire system & MDP & DDQN & Only for a singel MEC system model \\\midrule				
				\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
				\cite{wu2024privacy} 2& MEC-based IIoT  & Joint power allocation and computation offloading  & Optimization of privacy protection and quality of service. & MDP & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{zhao2019deep} & MEC & User association and resource allocation & Optimize downlink utility & MDP & Multi-agent D3QN & Not consider system delay and energy consumtion \\\midrule
				\cite{sun2024hierarchical} & MEC & Computation offloading and service caching & Minimize average service delay & MDP & hierarchical DRL & Not consider system energy consumtion \\\midrule
				\cite{wu2023computation} 1 & SDN-based MEC & Stochastic game-based resource allocation & Minimize energy consumption and processing delay & MDP & Multi-agent RL & Not consider tasks with maximum delay tolerance\\\midrule
				\cite{dai2020edge} & MEC & Resource allocation and computation offloading & Minimize energy consumption & MDP & DDPG & Not take user's demand into consideration \\\midrule
				\cite{liu2021learn} & MEC & Resource allocation and computation offloading & Minimize execution cost & POMDP & Multi-agent actor-critic network &  Not consider delay-sensitive tasks \\\midrule
				\cite{gong2022edge} & MEC-based IIoT & Resource allocation and computation offloading & Minimize task computation delay and energy consumption & MDP & DRL & Not take user's demand into consideration\\\midrule
				\cite{gao2022large} + & Queuing-based heterogeneous MEC & Decentralized computation offloading & Optimize system cost and completion rates & MDP & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{9253665} + &  Queuing-based MEC &  Computation offloading & Minimize drop rate and computation delay. & MDP & D3QN + LSTM & Not consider system energy consumtion\\\midrule
				\cite{wu2024combining} 1 & Queuing-based IIoT MEC & Privacy aware computation offloading & Optimization of computation rate and energy consumption & MDP & Actor-critic network  & Not take user's demand into consideration \\\midrule
				\cite{wu2023multi} 1& Queuing-based IIoT MEC & Computation offloading & Optimize delay and energy consumption & POMDP & Multi-agent DQN&  Not consider tasks with maximum delay tolerance \\\midrule
				\cite{huang2021deadline} & Queuing-based MEC & Computation offloading & Optimize drop rate and energy consumption & POMDP & DDPG & Not consider computation delay \\ \midrule
				\cite{Bolourian-WCL24} - & Queuing-based MEC with EHs& Energy harvesting  computation offloading & Minimize drop rate and energy consumption. & MDP  & DQN & Not consider computation delay  \\\midrule
				QECO &  Queuing-based MEC &  Computation offloading & Maximize the QoE of each MD individually &MDP & D3QN + LSTM & \\
				\toprule
		\end{tabular}}
		\label{table1}}
\end{table*}

  \textcolor{blue}{To reflecting real-world scenarios more accurately, the authors applied queuing theory to model MEC systems. It helps in modeling the dynamic nature of network, allowing MEC systems to allocate computational and transmission resources more efficiently.
 	To minimize energy consumption, Munir \textit{et al.} \cite{munir2021multi} developed a semi-distributed approach using a multi-agent RL framework for self-powered MEC. 
 	Liao \textit{et al.} in \cite{liao2023online} introduced a double DRL algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To investigate privacy aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} proposed a DQN-based method. They transform the problem on MDP to optimize computation rate and energy consumption in a queuing-based IIoT network.
	In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. 
	 Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks.
	 In above research \cite{munir2021multi}--\cite{huang2021deadline}, while queuing models were employed in several parts of the system, the offloading process was not approached as a holistic queuing system. }

@@

\subsection{Simulation Setup}
We consider a MEC environment with 50 MDs and 5 ENs, similar to \cite{9253665}. We also follow the model presented in \cite{zhou2021deep} to determine the energy consumption. All the parameters are given in Table~\ref{table}. To train the MDs' neural networks, we adopt a scenario comprising 1000 episodes. Each episode contains 100 time slots, each of length 0.1 second. The QECO algorithm incorporates real-time experience into its training process to continuously enhance the offloading strategy. Specifically, we employ a batch size of 16, maintain a fixed learning rate of 0.001, and set the discount factor $\gamma$ to 0.9. The probability of random exploration gradually decreases from an initial value 1, progressively approaching 0.01, all of which is facilitated by an RMSProp optimizer. \textcolor{blue}{Detailed settings regarding the neural networks and simulation can be found at our github repository.\footnote{\textcolor{blue}{The code for reproducing the simulation results of this letter is available at: \href{https://github.com/ImanRHT/QECO}{https://github.com/ImanRHT/QECO.}}}}

@@

\textcolor{blue}{We finally delve into the investigation of the convergence performance of the QECO algorithm in Fig.~\ref{chart01} and Fig.~\ref{chart02}. To validate the effectiveness of the QECO algorithm, we assess its convergence speed and rate compared to the vanilla DQN and D3QN-based computation offloading methods, measured by the average QoE across episodes. As shown in Fig.~\ref{chart01}, the average QoE for all computation offloading algorithms rises consistently as the episodes progress, until reaching convergence. Specifically, the vanilla DQN, D3QN, and our proposed algorithm converge after approximately 650, 550, and 400 iterations, respectively. This demonstrates the relatively faster convergence of the QECO algorithm, which also achieves a higher average QoE than the other methods. These results highlight the positive impact of workload prediction by the LSTM network and its advantage in efficiently utilizing the processing capabilities of MDs and ENs.}

@@

\section{Conclusion}\label{section:VII}
%In this study, our focus was directed towards addressing the challenge of offloading in MEC systems, where strict task processing deadlines and energy constraints adversely impact system performance. To address this challenge, we propose the QECO algorithm specifically designed to empower MDs to make offloading decisions without relying on knowledge about task models or other MDs' offloading decisions. The QECO algorithm not only adapts to the uncertain dynamics of load levels at ENs but also effectively manages the ever-changing system environment. Through extensive simulations, we demonstrate that QECO outperforms several established benchmark techniques. Specifically, it can significantly increase the average user's QoE. This advantage can lead to improvements in key performance metrics, including task completion rate, task delay, and energy consumption, according to the user's demands under different system conditions.

>>>>

	\item \textit{Performance Evaluation:} \textcolor{blue}{We conduct comprehensive experiments to evaluate the QECO's performance under various computational workloads, as well as its training convergence with different techniques and hyperparameters. The results demonstrate that our algorithm effectively utilizes the computing resources of MDs and ENs, addressing the dynamic workloads at ENs. It converges more quickly compared to vanila DQN and DDQN methods, while providing a substantial improvement of at least 34.8\% in average QoE. This advantage is achieved through prioritizing MD's QoE preferences, resulting in improvements of at least 12.2\%, 11.7\%, and 10.8\% in the average number of completed tasks, task delay, and energy consumption, respectively, when compared to the distributed and collective DRL-based offloading algorithm (DCDRL) \cite{qiu2020distributed} and several benchmark methods.}

@@

	  \textcolor{blue}{Some studies explored task offloading problems in more advanced scenarios. Chen \textit{et al.} in \cite{chen2021drl} addressed a joint optimization problem involving computation offloading and resource allocation, aiming to reduce both task processing delay and energy consumption across all MDs.
	 To optimize privacy protection and quality of service, authors in \cite{wu2024privacy} investigated the joint computation offloading and power allocation problems for the IIoT network. They modeled the problem as an MDP and proposed a multi-agent deep Q-network (DQN)-based algorithm.
	 Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm based on deep deterministic policy gradient (DDPG) to concurrently optimize resource allocation and computation offloading.
	 In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
	 These works did not take into account the challenges of a dynamic network environment. For instance, in \cite{chen2021drl}, MDs are assumed to have sufficient transmission and computation resources from ENs, which may not always align with real-world conditions. While the works of \cite{wu2024privacy}--\cite{zhao2019deep} investigated the limited resources in MEC, they did not consider the dynamic workload at ENs in decision making processes, imposing significant challenges in overall system performance. }

@@

	 \begin{table*}[htbp]\textcolor{blue}{
		\renewcommand{\arraystretch}{0.7}
		\captionsetup{name=TABLE}
		\caption{Comparison of Related Works}
		\scalebox{1}{%
			\begin{tabular}{ lp{2.2cm}p{3cm}p{3.4cm}p{0.7cm}p{2cm}p{3.2cm}l} 
				\toprule
				\textbf{Paper} & \textbf{Scenario} &  \textbf{Problem} &  \textbf{Objective} &  \textbf{Model} & \textbf{Method} &  \textbf{Drawbacks}  \\ \midrule
				\cite{liao2023online} & Queuing-based MEC & Online computation offloading & Minimize computation delay and energy consumption & MDP & DDQN & Only for a singel MEC system model  \\\midrule
				\cite{zhou2021deep} & heterogeneous MEC & Resource allocation and computation offloading & Minimize energy consumption of the entire system & MDP & DDQN & Only for a singel MEC system model \\\midrule				
				\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
				\cite{wu2024privacy} & MEC-based IIoT  & Joint power allocation and computation offloading  & Optimization of privacy protection and quality of service. & MDP & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{zhao2019deep} & MEC & User association and resource allocation & Optimize downlink utility & MDP & Multi-agent D3QN & Not consider system delay and energy consumtion \\\midrule
				\cite{sun2024hierarchical} & MEC & Computation offloading and service caching & Minimize average service delay & MDP & hierarchical DRL & Not consider system energy consumtion \\\midrule
				\cite{wu2023computation}  & SDN-based MEC & Stochastic game-based resource allocation & Minimize energy consumption and processing delay & MDP & Multi-agent RL & Not consider tasks with maximum delay tolerance\\\midrule
				\cite{dai2020edge} & MEC & Resource allocation and computation offloading & Minimize energy consumption & MDP & DDPG & Not take user's demand into consideration \\\midrule
				\cite{liu2021learn} & MEC & Resource allocation and computation offloading & Minimize execution cost & POMDP & Multi-agent actor-critic network &  Not consider delay-sensitive tasks \\\midrule
				\cite{gong2022edge} & MEC-based IIoT & Resource allocation and computation offloading & Minimize task computation delay and energy consumption & MDP & DRL & Not take user's demand into consideration\\\midrule
				\cite{gao2022large} & Queuing-based heterogeneous MEC & Decentralized computation offloading & Optimize system cost and completion rates & MDP & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{9253665} &  Queuing-based MEC &  Computation offloading & Minimize drop rate and computation delay. & MDP & D3QN + LSTM & Not consider system energy consumtion\\\midrule
				\cite{wu2024combining}  & Queuing-based IIoT MEC & Privacy aware computation offloading & Optimization of computation rate and energy consumption & MDP & Actor-critic network  & Not take user's demand into consideration \\\midrule
				\cite{wu2023multi} 1& Queuing-based IIoT MEC & Computation offloading & Optimize delay and energy consumption & POMDP & Multi-agent DQN&  Not consider tasks with maximum delay tolerance \\\midrule
				\cite{huang2021deadline} & Queuing-based MEC & Computation offloading & Optimize drop rate and energy consumption & POMDP & DDPG & Not consider computation delay \\ \midrule
				\cite{Bolourian-WCL24}  & Queuing-based MEC with EHs& Energy harvesting  computation offloading & Minimize drop rate and energy consumption. & MDP  & DQN & Not consider computation delay  \\\midrule
				QECO &  Queuing-based MEC &  Computation offloading & Maximize the QoE of each MD individually &MDP & D3QN + LSTM & \\
				\toprule
		\end{tabular}}
		\label{table1}}
\end{table*}

  \textcolor{blue}{To reflecting real-world scenarios more accurately, the authors applied queuing theory to model MEC systems. It helps in modeling the dynamic nature of network, allowing MEC systems to allocate computational and transmission resources more efficiently.
 	To minimize energy consumption, Munir \textit{et al.} \cite{munir2021multi} developed a semi-distributed approach using a multi-agent RL framework for self-powered MEC. 
 	Liao \textit{et al.} in \cite{liao2023online} introduced a double DQN (DDQN)-based algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To investigate privacy aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} proposed a DQN-based method. They transform the problem on MDP to optimize computation rate and energy consumption in a queuing-based IIoT network.
	In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. 
	 Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks.
	 In above research \cite{munir2021multi}--\cite{huang2021deadline}, while queuing models were employed in several parts of the system, the offloading process was not approached as a holistic queuing system. }

@@

\subsection{Simulation Setup}
We consider a MEC environment with 50 MDs and 5 ENs, similar to \cite{9253665}. We also follow the model presented in \cite{zhou2021deep} to determine the energy consumption. All the parameters are given in Table~\ref{table}. To train the MDs' neural networks, we adopt a scenario comprising 1000 episodes. Each episode contains 100 time slots, each of length 0.1 second. The QECO algorithm incorporates real-time experience into its training process to continuously enhance the offloading strategy. Specifically, we employ a batch size of 16, maintain a fixed learning rate of 0.001, and set the discount factor $\gamma$ to 0.9. The probability of random exploration gradually decreases from an initial value 1, progressively approaching 0.01, all of which is facilitated by an RMSProp optimizer. \textcolor{blue}{The algorithm source code is publicly available in \cite{QECO}}.

@@

\textcolor{blue}{	
	We finally delve into the investigation of the convergence performance of the QECO algorithm in Fig.~\ref{chart01} and Fig.~\ref{chart02}. To validate the effectiveness of the QECO algorithm, we assess its convergence rate compared to the vanilla DQN and DDQN-based computation offloading methods, measured by the average QoE across episodes. As shown in Fig.~\ref{chart01}, the average QoE for all computation offloading algorithms rises consistently as the episodes progress, until reaching convergence. Specifically, the DQN, DDQN, and our proposed algorithm converge after approximately 650, 550, and 450 iterations, respectively. This demonstrates the relatively faster convergence of the QECO algorithm, which also achieves a higher average QoE than the other methods. These results highlight the positive impact of workload prediction by the LSTM network and its advantage in efficiently utilizing the processing capabilities of MDs and ENs.}

@@

\section{Conclusion}\label{section:VII}

%In this study, our focus was directed towards addressing the challenge of offloading in MEC systems, where strict task processing deadlines and energy constraints adversely impact system performance. To address this challenge, we propose the QECO algorithm specifically designed to empower MDs to make offloading decisions without relying on knowledge about task models or other MDs' offloading decisions. The QECO algorithm not only adapts to the uncertain dynamics of load levels at ENs but also effectively manages the ever-changing system environment. Through extensive simulations, we demonstrate that QECO outperforms several established benchmark techniques. Specifically, it can significantly increase the average user's QoE. This advantage can lead to improvements in key performance metrics, including task completion rate, task delay, and energy consumption, according to the user's demands under different system conditions.
